---
title: "FYP Team Data"
author: "Cian Hughes"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(tidyr)
library(ggsoccer)

```

```{r}
# Manually set colors for each type_name
color_mapping <- c(
  corner_crossed = "pink",
  dribble = "yellow",
  pass = "red",
  shot = "darkblue",
  corner_short = "orange",
  goalkick = "black",
  cross = "darkgreen",
  freekick_crossed = "lightgreen",
  take_on = "brown",
  interception = "violet",
  shot_freekick = "salmon"
)
```

```{r}
# Define the custom soccer pitch with scaled y-axis
# Pitch in Metres is 105 x 68
# Halfway = 52.5 on the x-axis
pitch_custom <- list(
  length = 105,
  width = 68,
  penalty_box_length = 16,
  penalty_box_width = 43,
  six_yard_box_length = 6,
  six_yard_box_width = 18,
  penalty_spot_distance = 10,
  goal_width = 8,
  origin_x = 0,
  origin_y = 0
)
```

```{r}
# Define the coordinates for your custom boxes
custom_zones <- list(
  zone1 = list(x_min = 0, x_max = 30, y_min = 0, y_max = 68),
  zone2 = list(x_min = 30, x_max = 52.5, y_min = 0, y_max = 68),
  zone3 = list(x_min = 52.5, x_max = 75, y_min = 0, y_max = 68),
  zone4 = list(x_min = 75, x_max = 105, y_min = 0, y_max = 68)
)

# Create a data frame from the custom boxes list
custom_zones_df <- data.frame(
  zone = names(custom_zones),
  x = sapply(custom_zones, function(zone) mean(c(zone$x_min, zone$x_max))),
  y = sapply(custom_zones, function(zone) mean(c(zone$y_min, zone$y_max))),
  x_min = sapply(custom_zones, function(zone) zone$x_min),
  x_max = sapply(custom_zones, function(zone) zone$x_max),
  y_min = sapply(custom_zones, function(zone) zone$y_min),
  y_max = sapply(custom_zones, function(zone) zone$y_max)
)

# Function to determine the box for each point
find_zone <- function(x, y) {
  zone_names <- sapply(seq_along(x), function(i) {
    for (zone_name in names(custom_zones)) {
      zone <- custom_zones[[zone_name]]
      if (x[i] >= zone$x_min && x[i] <= zone$x_max && y[i] >= zone$y_min && y[i] <= zone$y_max) {
        return(zone_name)
      }
    }
    return(NA)
  })
  return(zone_names)
}
```

```{r}

ggplot() +
  annotate_pitch(dimensions = pitch_custom) +
  theme_pitch() +
  geom_rect(data = custom_zones_df, aes(xmin = x_min, xmax = x_max, ymin = y_min, ymax = y_max), fill = NA, color = "black") +
  geom_text(data = custom_zones_df, aes(x = x, y = y, label = zone), size = 4, color = "black") +
  labs(title = "Zones on Pitch") +
  theme_minimal()

```
# Now getting all sequences for any team_id

```{r}
# Read the "actions" CSV file
actions <- read.csv("data/actions.csv") # Change path file if needed

# Read the "players" CSV file
players <- read.csv("data/players.csv") # Change path file if needed

# Create a new column in the data for player names
actions$player_name <- players$shortName[match(actions$player_id, players$wyId)]

```

# New code for extracting one team's sequences starts here:
```{r}
# Function to filter data for a specific team ID
filter_data_for_team <- function(actions, target_team_id) {
  game_ids_with_team <- unique(actions$game_id[actions$team_id == target_team_id])
  
  filtered_data <- actions[actions$game_id %in% game_ids_with_team, ]
  
  return(filtered_data)
}

desired_team_id <- 1625  # Replace with your desired team ID

filtered_data <- filter_data_for_team(actions, desired_team_id)

```

# Function that will return all sequences for the given team_id

```{r}
# Function to extract all sequences containing a specific team_id
extract_all_sequences <- function(data, target_team_id) {
  sequences <- list()

  count <- 0

  target_in_sequence <- FALSE

  for (i in seq_len(nrow(data))) {
    if (i == 1 || data$team_id[i] != data$team_id[i - 1]) {
      count <- count + 1

      current_sequence <- data[i, , drop = FALSE]

      if (data$team_id[i] == target_team_id) {
        target_in_sequence <- TRUE
      }
    } else {
      current_sequence <- rbind(current_sequence, data[i, , drop = FALSE])

      if (data$team_id[i] == target_team_id) {
        target_in_sequence <- TRUE
      }
    }

    if (i == nrow(data) || data$team_id[i + 1] != data$team_id[i]) {
      if (target_in_sequence || (nrow(current_sequence) == 1 && current_sequence$team_id == target_team_id)) {
        if (!any(grepl("^corner_crossed|^corner_short|^freekick_crossed|^freekick_short|^throw_in", current_sequence$type_name)) && nrow(current_sequence) > 1) {
          sequences[[count]] <- current_sequence
        }
      }

      current_sequence <- NULL  
      target_in_sequence <- FALSE 
    }
  }

  return(sequences)
}

# Call the function and store the results
result_all_sequences <- extract_all_sequences(filtered_data, target_team_id = desired_team_id)

```


# Filtering for the starting points of each sequence
```{r}
# Extract only the starting point of each sequence
all_starting_points <- lapply(result_all_sequences, function(seq) seq[1, ])

# Convert starting points to a data frame
all_starting_points_df <- do.call(rbind, all_starting_points)

```

# Display all these on a pitch with the zones, and a histogram showing the count of points per zone
```{r}

all_starting_points_df$zone <- find_zone(all_starting_points_df$start_x, all_starting_points_df$start_y)

ggplot() +
  annotate_pitch(dimensions = pitch_custom) +
  theme_pitch() +
  geom_rect(data = custom_zones_df, aes(xmin = x_min, xmax = x_max, ymin = y_min, ymax = y_max), fill = NA, color = "black") +
  geom_point(data = all_starting_points_df, aes(x = start_x, y = start_y, color = zone), size = 2) +
  labs(title = paste("Starting Points of Sequences")) +
  scale_color_manual(values = c("red", "blue", "green", "cyan", "yellow", "black", "salmon", "hotpink","beige")) +
  theme_minimal()

all_zone_counts <- table(factor(all_starting_points_df$zone, levels = names(custom_zones)))

all_zone_counts_df <- data.frame(zone = names(all_zone_counts), count = as.numeric(all_zone_counts))

all_zone_counts_df$color <- c("red", "blue", "green", "cyan", "yellow", "black", "salmon", "hotpink","beige")[match(all_zone_counts_df$zone, names(custom_zones))]

ggplot(all_zone_counts_df, aes(x = zone, y = count, fill = color)) +
  geom_bar(stat = "identity", color = "black") +
  geom_text(aes(label = count), vjust = -0.5, color = "black", size = 3) +
  labs(title = "Number of Points in Each Zone (All Sequences)", x = "zone", y = "Count") +
  scale_fill_identity() +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

# Now do the same but for sequences that contain a shot action

```{r}
# Function to extract all sequences containing a specific type and team_id
extract_all_shots <- function(data, target_type = "shot", target_team_id) {
  sequences <- list()
  count <- 0
  
  target_in_sequence <- FALSE
  
  for (i in seq_len(nrow(data))) {
    if (i == 1 || data$team_id[i] != data$team_id[i - 1]) {
      count <- count + 1
      
      current_sequence <- data[i, , drop = FALSE]
      
      if (data$type_name[i] == target_type && data$team_id[i] == target_team_id) {
        target_in_sequence <- TRUE
      }
    } else {
      current_sequence <- rbind(current_sequence, data[i, , drop = FALSE])
      
      if (data$type_name[i] == target_type && data$team_id[i] == target_team_id) {
        target_in_sequence <- TRUE
      }
    }
    
    if (i == nrow(data) || data$team_id[i + 1] != data$team_id[i]) {
      if (target_in_sequence || (nrow(current_sequence) == 1 && current_sequence$type_name == target_type && current_sequence$team_id == target_team_id)) {
        if (!any(grepl("^corner_crossed|^corner_short|^freekick_crossed|^freekick_short|^throw_in", current_sequence$type_name)) && nrow(current_sequence) > 1) {
          sequences[[count]] <- current_sequence
        }
      }
      
      current_sequence <- NULL  
      target_in_sequence <- FALSE 
    }
  }
  
  return(sequences)
}

target_types <- c("shot")
result_all_shots <- list()

for (target_type in target_types) {
  result_all_shots <- c(result_all_shots, extract_all_shots(filtered_data, target_type = target_type, target_team_id = desired_team_id))
}

```

# Then filter again for the starting points
# Then plot all on a graph and display the count

```{r}
# Extract only the starting point of each sequence
shot_starting_points <- lapply(result_all_shots, function(seq) seq[1, ])

# Convert starting points to a data frame
shot_starting_points_df <- do.call(rbind, shot_starting_points)

```

# Display all these on a pitch with the zones, and a histogram showing the count of points per zone
```{r}

shot_starting_points_df$zone <- find_zone(shot_starting_points_df$start_x, shot_starting_points_df$start_y)

ggplot() +
  annotate_pitch(dimensions = pitch_custom) +
  theme_pitch() +
  geom_rect(data = custom_zones_df, aes(xmin = x_min, xmax = x_max, ymin = y_min, ymax = y_max), fill = NA, color = "black") +
  geom_point(data = shot_starting_points_df, aes(x = start_x, y = start_y, color = zone), size = 2) +
  labs(title = paste("Starting Points of Sequences Containing a shot by team",desired_team_id)) +
  scale_color_manual(values = c("red", "blue", "green", "cyan", "yellow", "black", "salmon", "hotpink","beige")) +
  theme_minimal()

shot_zone_counts <- table(factor(shot_starting_points_df$zone, levels = names(custom_zones)))

shot_zone_counts_df <- data.frame(zone = names(shot_zone_counts), count = as.numeric(shot_zone_counts))

shot_zone_counts_df$color <- c("red", "blue", "green", "cyan", "yellow", "black", "salmon", "hotpink","beige")[match(shot_zone_counts_df$zone, names(custom_zones))]

ggplot(shot_zone_counts_df, aes(x = zone, y = count, fill = color)) +
  geom_bar(stat = "identity", color = "black") +
  geom_text(aes(label = count), vjust = -0.5, color = "black", size = 3) +
  labs(title = "Number of Points in Each Zone (Sequences Containing a Shot)", x = "zone", y = "Count") +
  scale_fill_identity() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

# Now I want to display 
# 1. Graph with all sequences on the pitch
# 2. Graph with all shot sequences
# 3. Graph that gives the boxes with each box labelled with the ratio of sequences to shots

```{r}
# 1.
all_starting_points_df$zone <- find_zone(all_starting_points_df$start_x, all_starting_points_df$start_y)

ggplot() +
  annotate_pitch(dimensions = pitch_custom) +
  theme_pitch() +
  geom_rect(data = custom_zones_df, aes(xmin = x_min, xmax = x_max, ymin = y_min, ymax = y_max), fill = NA, color = "black") +
  geom_point(data = all_starting_points_df, aes(x = start_x, y = start_y, color = zone), size = 2) +
  labs(title = paste("Starting Points of Sequences By Team",desired_team_id)) +
  scale_color_manual(values = c("red", "blue", "green", "cyan", "yellow", "black", "salmon", "hotpink","beige")) +
  theme_minimal()

# 2.
shot_starting_points_df$zone <- find_zone(shot_starting_points_df$start_x, shot_starting_points_df$start_y)

ggplot() +
  annotate_pitch(dimensions = pitch_custom) +
  theme_pitch() +
  geom_rect(data = custom_zones_df, aes(xmin = x_min, xmax = x_max, ymin = y_min, ymax = y_max), fill = NA, color = "black") +
  geom_point(data = shot_starting_points_df, aes(x = start_x, y = start_y, color = zone), size = 2) +
  labs(title = paste("Starting Points of Sequences Containing a Shot By Team",desired_team_id)) +
  scale_color_manual(values = c("red", "blue", "green", "cyan", "yellow", "black", "salmon", "hotpink","beige")) +
  theme_minimal()

# 3.
shot_percentage <- shot_zone_counts_df$count / all_zone_counts_df$count * 100

custom_zones_df$percentage <- round(shot_percentage, 2)

ggplot() +
  annotate_pitch(dimensions = pitch_custom) +
  theme_pitch() +
  geom_rect(data = custom_zones_df, aes(xmin = x_min, xmax = x_max, ymin = y_min, ymax = y_max), fill = NA, color = "black") +
  geom_text(data = custom_zones_df, aes(x = x, y = y, label = paste0(percentage, "%")), size = 4, color = "black") +
  labs(title = "Percentage of Sequences That Result in a Shot Per Zone") +
  theme_minimal()
```
# Now filter for goals - aka successful shots

```{r}
goal_sequences <- function(sequences) {
  filtered_sequences <- list()
  
  for (i in seq_along(sequences)) {
    # Check if the sequence contains the specified target_type with result_id "1" (indicating success)
    # and exclude sequences with type_name starting with specified values
    if (any(sequences[[i]]$result_id == "1" & 
            sequences[[i]]$type_name %in% c("shot")) &&
        !any(grepl("^corner_crossed|^corner_short|^freekick_crossed|^freekick_short|^throw_in", sequences[[i]]$type_name))) {
      filtered_sequences <- c(filtered_sequences, list(sequences[[i]]))
    }
  }
  
  return(filtered_sequences)
}

goal_zones <- goal_sequences(result_all_shots)

```


```{r}
# Extract only the starting point of each sequence in goal sequences
goal_starting_points <- lapply(goal_zones, function(seq) seq[1, ])

goal_starting_points_df <- do.call(rbind, goal_starting_points)

goal_starting_points_df$zone <- find_zone(goal_starting_points_df$start_x, goal_starting_points_df$start_y)

ggplot() +
  annotate_pitch(dimensions = pitch_custom) +
  theme_pitch() +
  geom_rect(data = custom_zones_df, aes(xmin = x_min, xmax = x_max, ymin = y_min, ymax = y_max), fill = NA, color = "black") +
  geom_point(data = goal_starting_points_df, aes(x = start_x, y = start_y, color = zone), size = 2) +
  labs(title = paste("Starting Points of Sequences Resulting in a Goal by team",desired_team_id)) +
  scale_color_manual(values = c("red", "blue", "green", "cyan", "yellow", "black", "salmon", "hotpink","beige")) +
  theme_minimal()

goal_zone_counts <- table(factor(goal_starting_points_df$zone, levels = names(custom_zones)))

goal_zone_counts_df <- data.frame(zone = names(goal_zone_counts), count = as.numeric(goal_zone_counts))

goal_zone_counts_df$color <- c("red", "blue", "green", "cyan", "yellow", "black", "salmon", "hotpink","beige")[match(goal_zone_counts_df$zone, names(custom_zones))]

ggplot(goal_zone_counts_df, aes(x = zone, y = count, fill = color)) +
  geom_bar(stat = "identity", color = "black") +
  geom_text(aes(label = count), vjust = -0.5, color = "black", size = 3) +
  labs(title = "Number of Points in Each Zone (Sequences Ending in a goal)", x = "zone", y = "Count") +
  scale_fill_identity() +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

```
```{r}
# 1.
shot_starting_points_df$zone <- find_zone(shot_starting_points_df$start_x, shot_starting_points_df$start_y)

ggplot() +
  annotate_pitch(dimensions = pitch_custom) +
  theme_pitch() +
  geom_rect(data = custom_zones_df, aes(xmin = x_min, xmax = x_max, ymin = y_min, ymax = y_max), fill = NA, color = "black") +
  geom_point(data = shot_starting_points_df, aes(x = start_x, y = start_y, color = zone), size = 2) +
  labs(title = paste("Starting Points of Sequences Containing a Shot By Team",desired_team_id)) +
  scale_color_manual(values = c("red", "blue", "green", "cyan", "yellow", "black", "salmon", "hotpink","beige")) +
  theme_minimal()

# 2.
goal_starting_points_df$zone <- find_zone(goal_starting_points_df$start_x, goal_starting_points_df$start_y)

ggplot() +
  annotate_pitch(dimensions = pitch_custom) +
  theme_pitch() +
  geom_rect(data = custom_zones_df, aes(xmin = x_min, xmax = x_max, ymin = y_min, ymax = y_max), fill = NA, color = "black") +
  geom_point(data = goal_starting_points_df, aes(x = start_x, y = start_y, color = zone), size = 2) +
  labs(title = paste("Starting Points of Sequences Resulting in a Goal By Team",desired_team_id)) +
  scale_color_manual(values = c("red", "blue", "green", "cyan", "yellow", "black", "salmon", "hotpink","beige")) +
  theme_minimal()

# 3.
# Need to find the % of shots to sequence
# Calculate the percentage of sequences containing a shot action for each zone
shot_percentage <- goal_zone_counts_df$count / shot_zone_counts_df$count * 100

custom_zones_df$percentage <- round(shot_percentage, 2)

ggplot() +
  annotate_pitch(dimensions = pitch_custom) +
  theme_pitch() +
  geom_rect(data = custom_zones_df, aes(xmin = x_min, xmax = x_max, ymin = y_min, ymax = y_max), fill = NA, color = "black") +
  geom_text(data = custom_zones_df, aes(x = x, y = y, label = paste0(percentage, "%")), size = 4, color = "black") +
  labs(title = "Percentage of Shot Sequences That Result in a Goal Per Zone") +
  theme_minimal()

# 4.
# Need to find the % of goals to sequences
goal_percentage <- goal_zone_counts_df$count / all_zone_counts_df$count * 100

custom_zones_df$percentage <- round(goal_percentage, 2)

ggplot() +
  annotate_pitch(dimensions = pitch_custom) +
  theme_pitch() +
  geom_rect(data = custom_zones_df, aes(xmin = x_min, xmax = x_max, ymin = y_min, ymax = y_max), fill = NA, color = "black") +
  geom_text(data = custom_zones_df, aes(x = x, y = y, label = paste0(percentage, "%")), size = 4, color = "black") +
  labs(title = "Percentage of Sequences That Result in a Goal Per Zone") +
  theme_minimal()
```
